---
title: "shiny_harvesting"
author: "Irantzu Lamarca"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DATA HARVESTING

## Web Scraping

We will start by performing the web scraping from the ATP Tour webpage in order to obtain data from the ATP raking tennis players. 

```{r}
# Libraries
library(scrapex)
library(rvest)
library(httr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(xml2)
library(lubridate)
```

```{r}

atp_link <- "https://www.atptour.com/es/rankings/singles?rankRange=0-100&region=all&dateWeek=Semana%20Actual&SortField=Ranking&SortAscending=True"
browseURL(prep_browser(atp_link))


# the website does not let you scrape it unless an user-agent is set.

html_atp_website <- httr::GET(atp_link, 
                          httr::add_headers("User-Agent" = "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/134.0.0.0 Safari/537.36 Edg/134.0.0.0
; Diego Fernandez / dieguitus23@gmail.com")) |>
  read_html()


#just change the user agent that is in the script into the user agent that you have
```

### Ranking tables

We will try to retrieve the ranking tables first

```{r}
# let´s try to retrieve the tables: there are only 2 tables and they seem the same, I´ll extract the 1st one.
# It looks like we can only extract in the tables the players that did not move up or down in the ranking.

all_tables <- html_atp_website |> html_table()

atp_table <- all_tables[[1]]
atp_table


# let´s try another way: Regular expressions: the // looks for every "span" and [@class='lastName'] means: 
# Select elements that have an attribute class with the value "lastName"
# xml_text only extracts the text part, the name of the players.

player_name <- xml_find_all(html_atp_website, "//span[@class='lastName']") |> 
  xml_text()
player_name

# let´s store it in a table:
atp_ranking <- tibble(
  ranking = 1:100,
  name = player_name)
atp_ranking
# Let´s now extract the next 100 players:
atp_link2 <-"https://www.atptour.com/es/rankings/singles?rankRange=101-200&region=all&dateWeek=Semana+Actual&SortField=Ranking&SortAscending=True"

html_atp_website2 <- httr::GET(atp_link2, 
                              httr::add_headers("User-Agent" = "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/134.0.0.0 Safari/537.36 Edg/134.0.0.0
; Diego Fernandez / dieguitus23@gmail.com")) |>
  read_html()

player_name2 <- xml_find_all(html_atp_website2, "//span[@class='lastName']") |> 
  xml_text()
player_name2

# let´s store it in a table:
atp_ranking2 <- tibble(
  ranking = 101:200,
  name = player_name2)
atp_ranking2

# Let´s extract the last 100 players
atp_link3 <-"https://www.atptour.com/es/rankings/singles?rankRange=201-300&region=all&dateWeek=Semana+Actual&SortField=Ranking&SortAscending=True"

html_atp_website3 <- httr::GET(atp_link3, 
                               httr::add_headers("User-Agent" = "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/134.0.0.0 Safari/537.36 Edg/134.0.0.0
; Diego Fernandez / dieguitus23@gmail.com")) |>
  read_html()

player_name3 <- xml_find_all(html_atp_website3, "//span[@class='lastName']") |> 
  xml_text()
player_name3

# let´s store it in a table:
atp_ranking3 <- tibble(
  ranking = 201:300,
  name = player_name3)
atp_ranking3

# Finally, let´s join the 3 tables and keep only the last name (surname)

final_ranking_table <- rbind(atp_ranking, atp_ranking2, atp_ranking3)
final_ranking_table

final_ranking_table <- final_ranking_table |> 
  mutate(name = gsub("^\\s+|\\s+$", "", name)) |>   # clean extra spaces
  mutate(name = sub(".*\\s+", "", name))
final_ranking_table
```

### Age, country, height and weight

We will now try to extract the age, country, height and weight

```{r}
# Example with Sinner: extracting age, country, height and weight from his web. We have run into some problems 
# because the variables we want to extract are not visible. We suspect selenium is needed to read javascript code.

require(RSelenium)


remDr <- rsDriver(port = 4444L, browser = "chrome")
driver <- remDr$client
driver$navigate("https://www.atptour.com/es/players/jannik-sinner/s0ag/overview")

Sys.sleep(10)


ul_element_left <- driver$findElement(using = "xpath", value = "//ul[contains(@class, 'pd_left')]")

ul_element_right <- driver$findElement(using = "xpath", value = "//ul[contains(@class, 'pd_right')]")



li_elements_left <- ul_element_left$findElements(using = "xpath", value = ".//li")
li_elements_right <- ul_element_right$findElements(using = "xpath", value = ".//li")


# Get the text of all the elements <li> LEFT
li_texts <- sapply(li_elements_left, function(x) x$getElementText())
full_text <- paste(li_texts, collapse = " ")
full_text

# Get the text of all the elements <li> RIGHT

li_texts_right <- sapply(li_elements_right, function(x) x$getElementText())
full_text_right <- paste(li_texts_right, collapse = " ")

# Extracting specific values with regular expressions
country <- str_extract(full_text_right, "\\((.*?)\\)") |>  str_replace_all("[\\(\\)]", "")
age <- as.numeric(gsub(".*Edad\\s*(\\d+).*", "\\1", full_text))  # Extracts number 23 (Age)
weight <- as.numeric(gsub(".*Peso\\s*(\\d+).*", "\\1", full_text))  # Extracts number 77 (Weight)
height <- as.numeric(gsub(".*Estatura\\s*(\\d+).*", "\\1", full_text))  # Extracts number 191 (Height)

# Create a data frame with extracted values
df <- data.frame(age = age, weight = weight, height = height, country = country)
df

driver$close()
remDr$server$stop()

# We were able to scrape the variables we need for the Sinner example,
# now we just need to replicate that with every player.
# (doesn´t work because the app is continously blocking us, we will try to find another way of doing it)
```

### Wins and looses

Let's get the wins and looses 

```{r}
# Now let's get the wins/looses in 2025
library(RSelenium)
remDr <- rsDriver(port = 4444L, browser = "chrome")
driver <- remDr$client
driver$navigate("https://www.atptour.com/es/stats/win-loss-index?indexCategory=all&timeCategory=ytd&country=all&sortBy=index&sortDirection=desc")

Sys.sleep(5)
page_source <- driver$getPageSource()[[1]]

html_content <- read_html(page_source)

rows <- html_content %>%
  html_nodes("tr")

# Loop through rows and extract each cell’s content
table_data <- lapply(rows, function(row) {
  row %>%
    html_nodes("td") %>%
    html_text(trim = TRUE)
})

table_winloss <- do.call(rbind, lapply(table_data, function(x) {
  if (length(x) == 4) x else c(x, rep("", 4 - length(x)))  
})) %>% as.data.frame()


colnames(table_winloss) <- c("Player", "Index", "Titles", "WinLoss")
table_winloss

# let´s clean the table:
table_winloss <- table_winloss |> 
  # Remove the number prefix from the Player column
  mutate(Player = gsub("^\\d+", "", Player)) |> 
  mutate(Player = trimws(Player)) |>  # Clean any extra spaces
  
  # Split the WinLoss column into Wins and Losses
  separate(WinLoss, into = c("Wins", "Losses"), sep = "-", convert = TRUE) 

table_winloss <- table_winloss|> 
  slice(-1) # remove the first row
  

driver$close()
remDr$server$stop()
```

### Points won with first serve

Let's get the points won with the first serve

```{r}
# the points won with first serve
remDr <- rsDriver(port = 4444L, browser = "chrome")
driver <- remDr$client
driver$navigate("https://www.atptour.com/es/stats/individual-game-stats?factType=Aces&year=2024&surface=all&country=all&sortBy=percentage&sortDirection=desc")

Sys.sleep(5)
page_source <- driver$getPageSource()[[1]]

html_content <- read_html(page_source)

rows <- html_content %>%
  html_nodes("tr")

# Loop through rows and extract each cell’s content
table_data <- lapply(rows, function(row) {
  row %>%
    html_nodes("td") %>%
    html_text(trim = TRUE)
})

table_1st_serve <- do.call(rbind, lapply(table_data, function(x) {
  if (length(x) == 5) x else c(x, rep("", 5 - length(x)))  
})) %>% as.data.frame()


colnames(table_1st_serve) <- c("Player", "1st_serve_win_%", "Points_won", "Total_points", "Matches_played")
table_1st_serve


# Now, let´s clean the table:

table_1st_serve <- table_1st_serve |> 
  mutate(Player = gsub("^\\d+", "", Player)) |> 
  slice(-1)
table_1st_serve

driver$close()
remDr$server$stop()
```

### Aces

Let's get the aces

```{r}
# the aces

remDr <- rsDriver(port = 4444L, browser = "chrome")
driver <- remDr$client
driver$navigate("https://www.atptour.com/es/stats/individual-game-stats?factType=Aces&year=2024&surface=all&country=all&sortBy=percentage&sortDirection=desc")

Sys.sleep(5)

page_source <- driver$getPageSource()[[1]]

html_content <- read_html(page_source)

rows <- html_content %>%
  html_nodes("tr")

# Loop through rows and extract each cell’s content
table_data <- lapply(rows, function(row) {
  row %>%
    html_nodes("td") %>%
    html_text(trim = TRUE)
})

table_aces <- do.call(rbind, lapply(table_data, function(x) {
  if (length(x) == 3) x else c(x, rep("", 3 - length(x)))  
})) %>% as.data.frame()


colnames(table_aces) <- c("Player", "Aces_number", "Matches_played")
table_aces


# Now, let´s clean the table:

table_aces <- table_aces |> 
  mutate(Player = gsub("^\\d+", "", Player)) |> 
  slice(-1)
table_aces

driver$close()
remDr$server$stop()
```

### Games won % (serve)

Let's get the percentage of the wins of the player when he has to serve

```{r}
# games won % (when the player has to serve)

remDr <- rsDriver(port = 4444L, browser = "chrome")
driver <- remDr$client
driver$navigate("https://www.atptour.com/es/stats/individual-game-stats?factType=Service-Games-Won&year=2024&surface=all&country=all&sortBy=percentage&sortDirection=desc")

Sys.sleep(5)

page_source <- driver$getPageSource()[[1]]

html_content <- read_html(page_source)

rows <- html_content %>%
  html_nodes("tr")

# Loop through rows and extract each cell’s content
table_data <- lapply(rows, function(row) {
  row %>%
    html_nodes("td") %>%
    html_text(trim = TRUE)
})

table_games_won_serve <- do.call(rbind, lapply(table_data, function(x) {
  if (length(x) == 5) x else c(x, rep("", 5 - length(x)))  
})) %>% as.data.frame()


colnames(table_games_won_serve) <- c(
  "Player", "%games_won(serve)", "games_won_serving", 
  "total_games_serving", "Matches_played")
table_games_won_serve


# Now, let´s clean the table:

table_games_won_serve <- table_games_won_serve |> 
  mutate(Player = gsub("^\\d+", "", Player)) |> 
  slice(-1)

driver$close()
remDr$server$stop()
```

### Games won % (receive)

Let's get the percentage of the wins of the player when he has to receive the serve

```{r}
# games won % (when the player has to receive the serve)

remDr <- rsDriver(port = 4444L, browser = "chrome")
driver <- remDr$client
driver$navigate("https://www.atptour.com/es/stats/individual-game-stats?factType=Return-Games-Won&year=2024&surface=all&country=all&sortBy=percentage&sortDirection=desc")

Sys.sleep(5)

page_source <- driver$getPageSource()[[1]]

html_content <- read_html(page_source)

rows <- html_content %>%
  html_nodes("tr")

# Loop through rows and extract each cell’s content
table_data <- lapply(rows, function(row) {
  row %>%
    html_nodes("td") %>%
    html_text(trim = TRUE)
})

table_games_won_receiving <- do.call(rbind, lapply(table_data, function(x) {
  if (length(x) == 5) x else c(x, rep("", 5 - length(x)))  
})) %>% as.data.frame()


colnames(table_games_won_receiving) <- c(
  "Player", "%games_won(receive)", 
  "games_won_receiving", "total_games_receiving", "Matches_played")
table_games_won_receiving


# Now, let´s clean the table:

table_games_won_receiving <- table_games_won_receiving |> 
  mutate(Player = gsub("^\\d+", "", Player)) |> 
  slice(-1)

driver$close()
remDr$server$stop()
```

### % Break points saved

Let's get the percentage of the break points saved

```{r}
# break points saved % 

remDr <- rsDriver(port = 4444L, browser = "chrome")
driver <- remDr$client
driver$navigate("https://www.atptour.com/es/stats/individual-game-stats?factType=Break-Points-Saved&year=2024&surface=all&country=all&sortBy=percentage&sortDirection=desc")

Sys.sleep(5)

page_source <- driver$getPageSource()[[1]]

html_content <- read_html(page_source)

rows <- html_content |> 
  html_nodes("tr")

# Loop through rows and extract each cell’s content
table_data <- lapply(rows, function(row) {
  row %>%
    html_nodes("td") %>%
    html_text(trim = TRUE)
})

table_break_points_saved <- do.call(rbind, lapply(table_data, function(x) {
  if (length(x) == 5) x else c(x, rep("", 5 - length(x)))  
})) %>% as.data.frame()


colnames(table_break_points_saved) <- c(
  "Player", "%break_points_saved", 
  "points_won", "total_points", "Matches_played")
table_break_points_saved




# Now, let´s clean the table:

table_break_points_saved <- table_break_points_saved |> 
  mutate(Player = gsub("^\\d+", "", Player)) |> 
  slice(-1)

driver$close()
remDr$server$stop()
```

### Join tables

```{r}

# Finally, let´s join all the tables
All_player_variables <-table_break_points_saved |> 
  full_join(table_games_won_receiving, by = "Player") |> 
  full_join(table_games_won_serve, by = "Player") |> 
  full_join(table_aces, by = "Player") |> 
  full_join(table_1st_serve, by = "Player") |> 
  full_join(table_winloss, by = "Player")

All_player_variables

All_player_variables <- All_player_variables |> 
  mutate(Player = gsub("^\\s+|\\s+$", "", Player)) |>   # clean extra spaces
  mutate(Player = sub(".*\\s+", "", Player))

# Let´s join the All_player_variables with the final_ranking_table table.

players_table <- final_ranking_table |> left_join(All_player_variables, by = c("name" = "Player")) 
players_table


```

For the second part of the project we will just need and we will just use the data retrieved from before, so we will save this data in a csv in order to be able to work with it later (even if we are going to add some variables). 

```{r}
write.csv(players_table, "sss.csv", row.names = FALSE)
```

### Age + country

Now we will try to get the age and the country of the player from another different website as the previous one was not letting us do it

```{r}
# Let´s extract the age and country from another website: 

age_country_link <- "https://tennisabstract.com/reports/atpRankings.html" |>
  read_html()

age_country <- age_country_link |> html_table()

atp_table <- age_country[[1]]
atp_table

# Let´s clean the table:

atp_table <- atp_table[-1, ]
atp_table <- atp_table[-1, ]
atp_table <- atp_table[, 1:4]

colnames(atp_table) <- c("Rank", "Player", "Country", "Birthdate")

# Keep only the first 300 rows
atp_table <- atp_table |> slice(1:300)

# Calculate age from birthdate
atp_table <- atp_table |> 
  mutate(Birthdate = as.Date(Birthdate),
         Age = floor(interval(Birthdate, today()) / years(1)))


# Keep only Rank, Last_Name, and Age
atp_table <- atp_table |>  select(Rank, name, Age, Country)
atp_table$Rank <- as.numeric(atp_table$Rank)

# Finally, let´s join the atp_table and the players_table
players_table <- players_table |> left_join(atp_table, by = c("ranking" = "Rank"))


write.csv(players_table, "players_table.csv", row.names = FALSE)

View(players_table)

```

## Shiny app

After retrieving all that information, we decided to create 2 shiny apps in order to visualize the data retrieved and to make it useful


```{r}
# Libraries
library(tidyr)
library(shiny)
library(dplyr)
library(readr)
library(ggplot2)
library(stringr)
library(scales)
library(tidyverse)
library(shinythemes)
library(plotly)
library(fmsb)
library(ggplot2)
```

We will read the dataset that we saved before for the second part of the project

```{r}

df <- read_csv("sss.csv")

# Have a quick look at it 
head(df)

str(df)

summary(df)

View(df)

```

### Cleaning and transforming data 

- Remove duplicate or irrelevant columns
- Convert percentages to numeric values
- Handle missing values 
- Convert data into correct format

```{r}
# Rename columns to avoid spacing problems
colnames(df) <- gsub(" ", "_", colnames(df))

# CConvert percentages to numerical values
df$`%break_points_saved` <- as.numeric(gsub("%", "", df$`%break_points_saved`)) / 100
df$`%games_won(receive)` <- as.numeric(gsub("%", "", df$`%games_won(receive)`)) / 100

# Correct columns with commas 
df$total_games_receiving <- as.numeric(gsub(",", "", df$total_games_receiving))

# View number of NA values per column
na_counts <- colSums(is.na(df))
na_counts <- na_counts[na_counts > 0] 
print(na_counts)

# Display percentage of NA values per column
na_percentage <- (na_counts / nrow(df)) * 100
print(na_percentage)

# NA chart to visualize impact
na_data <- data.frame(Variable = names(na_percentage), Porcentaje = na_percentage)
ggplot(na_data, aes(x = reorder(Variable, -Porcentaje), y = Porcentaje)) +
  geom_bar(stat="identity", fill="red") +
  coord_flip() +
  ggtitle("Percentage of NA Values per Variable") +
  xlab("Variable") +
  ylab("Percentage of NA Values")
```


```{r}
# Delete “Matches_played” and “Total_points” columns because they have 100% of NA
df <- df %>% select(-Matches_played, -Total_points)

```

```{r}
# Merge column data into one column, taking non-NA values
df$matches_played <- coalesce(df$Matches_played.x, df$Matches_played.y, df$Matches_played.x.x, df$Matches_played.y.y)

# Delete original columns
df <- df %>% select(-Matches_played.x, -Matches_played.y, -Matches_played.x.x, -Matches_played.y.y)


```

```{r}
# Check for NA again and for the variables names
na_counts <- colSums(is.na(df))
na_counts <- na_counts[na_counts > 0] 
print(na_counts)

na_percentage <- (na_counts / nrow(df)) * 100
print(na_percentage)

na_data <- data.frame(Variable = names(na_percentage), Porcentaje = na_percentage)
ggplot(na_data, aes(x = reorder(Variable, -Porcentaje), y = Porcentaje)) +
  geom_bar(stat="identity", fill="red") +
  coord_flip() +
  ggtitle("Percentage of NA Values per Variable") +
  xlab("Variable") +
  ylab("Percentage of NA")

```


```{r}
# Rename column "Index"
df <- df %>% rename(Win_loss_index = Index)
# Delete unwanted columns
df <- df %>% select(-Points_won, -total_points, -points_won)

colnames(df)
```

Now that we have the variables of the dataset as we would like, we can start with the shiny app. We know that we have lots of NA, but doing an imputation doesn't make sense in this case because the results wouldn't be real, so we will just drop_na() and we will work with the players that have all of the data

### Shiny app 1

We are going to create 2 shiny apps. The first one serves as an ATP Player Comparator, allowing users to select a professional tennis player and compare their performance metrics with custom input values. The app provides an interactive analysis of key statistics, including aces, win percentage, service and return games won, titles, and matches played. By selecting a player from a dropdown menu and adjusting these metrics, users can visualize the differences between their chosen values and the real performance data of professional players. The app generates a detailed statistical table, a bar chart for graphical comparison, a numerical table displaying the differences in metrics, and a radar chart for a more intuitive visual representation. The purpose of this tool is to provide players, coaches, and analysts with a practical way to compare their performance to elite players, helping to identify strengths and areas for improvement. By integrating interactive and visually engaging elements, this app enhances the accessibility and interpretability of performance data in tennis.

```{r}

# Converting percentages to numerical values and cleaning data
players_data <- df %>%
  mutate(across(contains("%"), ~as.numeric(str_remove(.x, "%"))/100)) %>%
  mutate(across(where(is.character), as.factor)) %>%
  drop_na()

# If the data are not a data.frame, force conversion
if (!inherits(players_data, "data.frame")) {
  players_data <- as.data.frame(players_data)
}

# Rename columns to avoid problematic characters
players_data <- players_data %>%
  rename(
    Break_Points_Saved = `%break_points_saved`,
    Games_Won_Receive = `%games_won(receive)`,
    Games_Won_Serve = `%games_won(serve)`,
    First_Serve_Win = `1st_serve_win_%`,
    Win_Percentage = `Win_loss_index`,
    Titles = `Titles`,
    Matches_Played = `matches_played`
  )

# Define application UI
ui <- fluidPage(
  theme = shinytheme("flatly"),
  titlePanel("ATP Player Comparator"),
  sidebarLayout(
    sidebarPanel(
      selectInput("player_choice", "Select a player to compare:", choices = unique(players_data$name)),
      numericInput("aces", "Number of Aces:", value = 50, min = 0, max = 1000),
      numericInput("win_loss", "Winning percentage:", value = 50, min = 0, max = 100),
      numericInput("serve_win", "% Games won when Serving:", value = 70, min = 0, max = 100),
      numericInput("return_win", "% Games won on Receiving:", value = 30, min = 0, max = 100),
      numericInput("titles", "Titles:", value = 1, min = 0, max = 100),
      numericInput("matches_played", "Matches played:", value = 50, min = 0, max = 1000),
      actionButton("compare", "Compare", class = "btn-primary"),
      actionButton("reset", "Reset values", class = "btn-warning")
    ),
    
    mainPanel(
      tabsetPanel(
        tabPanel("Statistics", tableOutput("player_stats")),
        tabPanel("Graphical Comparison", plotlyOutput("comparison_plot")),
        tabPanel("Difference in Metrics", tableOutput("difference_table")),
        tabPanel("Radar Comparison", plotOutput("radar_chart"))
      )
    )
  )
)

# Define server logic
server <- function(input, output, session) {
  player_stats <- reactive({
    req(input$player_choice)
    selected_player <- players_data %>%
      filter(name == input$player_choice) %>%
      select(Aces_number, Win_Percentage, Games_Won_Serve, Games_Won_Receive, Titles, Matches_Played) %>%
      as.data.frame()
    return(selected_player)
  })

  observeEvent(input$reset, {
    updateNumericInput(session, "aces", value = 50)
    updateNumericInput(session, "win_loss", value = 50)
    updateNumericInput(session, "serve_win", value = 70)
    updateNumericInput(session, "return_win", value = 30)
    updateNumericInput(session, "titles", value = 1)
    updateNumericInput(session, "matches_played", value = 50)
  })

  output$player_stats <- renderTable({
    player_stats()
  })

  output$difference_table <- renderTable({
    req(player_stats())
    diff_df <- data.frame(
      Métrica = c("Aces", "% Winning", "% Serving", "% Receiving", "Titles", "Games played"),
      Usuario = c(input$aces, input$win_loss, input$serve_win, input$return_win, input$titles, input$matches_played),
      Jugador = as.numeric(player_stats()[1, ]),
      Diferencia = abs(c(input$aces, input$win_loss, input$serve_win, input$return_win, input$titles, input$matches_played) -
                         as.numeric(player_stats()[1, ]))
    )
    diff_df
  })

  output$comparison_plot <- renderPlotly({
    req(player_stats())
    comparison_df <- data.frame(
      Métrica = c("Aces", "% Winning", "% Serving", "% Receiving", "Titles", "Games played"),
      Usuario = c(input$aces, input$win_loss, input$serve_win, input$return_win, input$titles, input$matches_played),
      Jugador = as.numeric(player_stats()[1, ])
    )
    ggplotly(
      ggplot(comparison_df, aes(x = Métrica)) +
        geom_bar(aes(y = Usuario, fill = "User"), stat = "identity", position = "dodge", alpha = 0.7) +
        geom_bar(aes(y = Jugador, fill = "Player"), stat = "identity", position = "dodge", alpha = 0.7) +
        labs(title = paste("Comparison with", input$player_choice), y = "Value", x = "Metric") +
        theme_minimal() +
        scale_fill_manual(values = c("User" = "blue", "Player" = "red"))
    )
  })

  output$radar_chart <- renderPlot({
    req(player_stats())
    radar_data <- rbind(rep(100, 6), rep(0, 6),
                        c(input$aces, input$win_loss, input$serve_win, input$return_win, input$titles, input$matches_played),
                        as.numeric(player_stats()[1, ]))
    colnames(radar_data) <- c("Aces", "% Winning", "% Serving", "% Receiving", "Titles", "Games played")
    radar_data <- as.data.frame(radar_data)
    radarchart(radar_data, axistype = 1, pcol = c("blue", "red"), plwd = 2, plty = 1, title = "Metric comparison")
  })
}

shinyApp(ui, server)

```

### Shiny app 2

This second app allows users to simulate tennis matches between two selected players based on statistical data. Users can choose two players and modify parameters such as fatigue level and surface effects, which influence the probability calculations. The app offers two simulation modes: a Monte Carlo simulation that predicts match outcomes over 1,000 iterations and a set-based simulation that determines the winner set by set. The results are displayed in a text summary, a bar plot for match outcomes, and a table if the set simulation is activated. Additionally, the app includes a "Live Odds" feature, visualizing the evolving win probabilities of both players throughout the simulations. By incorporating fatigue effects and surface considerations, this simulator provides a more dynamic and realistic representation of tennis matches, making it a valuable tool for fans, analysts, and players who want to assess potential matchups and performance factors.

```{r}
# Creating UI
ui_2 <- fluidPage(
  theme = shinytheme("flatly"),
  titlePanel("ATP Match Simulator"),
  
  sidebarLayout(
    sidebarPanel(
      selectInput("player1", "Select player 1:", choices = unique(players_data$name)),
      selectInput("player2", "Select player 2:", choices = unique(players_data$name)),
      sliderInput("fatigue_factor", "Fatigue Factor (Impact on Performance):", 
                  min = 0, max = 1, value = 0.5, step = 0.1),
      checkboxInput("surface_effect", "Consider type of surface", value = FALSE),
      checkboxInput("simulate_sets", "Simulate match by sets", value = FALSE),
      actionButton("simulate", "Simulate Match", class = "btn-primary")
    ),
    
    mainPanel(
      tabsetPanel(
        tabPanel("Result", 
                 h3("Simulation result"),
                 verbatimTextOutput("match_result"),
                 tableOutput("match_sets"),
                 plotOutput("match_plot")
        ),
        tabPanel("Live Odds", 
                 h3("Live Odds"),
                 plotOutput("match_evolution")
        )
      )
    )
  )
)

# Server
server_2 <- function(input, output) {
  
  simulate_match <- eventReactive(input$simulate, {
    
    # Extract statistics of selected players
    p1_stats <- players_data %>% filter(name == input$player1)
    p2_stats <- players_data %>% filter(name == input$player2)
    
    required_columns <- c("Games_Won_Serve", "Games_Won_Receive", "Aces_number", "Win_Percentage", "Matches_Played")
    
    if (!all(required_columns %in% colnames(players_data))) {
      return("Error: Required columns not found in the data.")
    }
    
    # NA management
    p1_stats[required_columns] <- lapply(p1_stats[required_columns], function(x) ifelse(is.na(x), 0, x))
    p2_stats[required_columns] <- lapply(p2_stats[required_columns], function(x) ifelse(is.na(x), 0, x))
    
    # Fatigue adjustment
    p1_fatigue <- 1 - input$fatigue_factor
    p2_fatigue <- 1 - input$fatigue_factor
    
    p1_prob <- ((p1_stats$Games_Won_Serve + p1_stats$Games_Won_Receive) / 2) * p1_fatigue
    p2_prob <- ((p2_stats$Games_Won_Serve + p2_stats$Games_Won_Receive) / 2) * p2_fatigue
    
    if (length(p1_prob) == 0 || length(p2_prob) == 0) {
      return("Error: Probabilities cannot be calculated due to missing values in the statistics.")
    }
    
    if (input$simulate_sets) {
      # Simulation by sets
      set_results <- replicate(5, {
        p1_win <- rbinom(1, 1, p1_prob)
        p2_win <- 1 - p1_win
        c(p1_win, p2_win)
      })
      
      p1_sets <- sum(set_results[1, ])
      p2_sets <- sum(set_results[2, ])
      
      return(list(
        p1_sets = p1_sets,
        p2_sets = p2_sets,
        set_results = data.frame(
          Set = 1:5,
          Ganador = c(rep(input$player1, p1_sets), rep(input$player2, p2_sets))
        )
      ))
      
    } else {
      # Monte Carlo simulation
      sim_results <- replicate(1000, {
        p1_score <- rbinom(1, 1, p1_prob)
        p2_score <- 1 - p1_score
        c(p1_score, p2_score)
      })
      
      p1_wins <- sum(sim_results[1, ])
      p2_wins <- sum(sim_results[2, ])
      
      return(list(
        p1_wins = p1_wins,
        p2_wins = p2_wins,
        sim_evolution = data.frame(
          Simulación = 1:1000,
          Probabilidad = cumsum(sim_results[1, ]) / (1:1000),
          Jugador = rep(input$player1, 1000)
        ) %>%
          bind_rows(data.frame(
            Simulación = 1:1000,
            Probabilidad = cumsum(sim_results[2, ]) / (1:1000),
            Jugador = rep(input$player2, 1000)
          ))
      ))
    }
  })

  # Show text result
  output$match_result <- renderPrint({
    result <- simulate_match()
    
    if (is.character(result)) {
      return(result)
    }
    
    if (input$simulate_sets) {
      p1_name <- input$player1
      p2_name <- input$player2
      paste(
        sprintf("%s ganó %d sets.", p1_name, result$p1_sets),
        sprintf("%s ganó %d sets.", p2_name, result$p2_sets),
        sep = "\n"
      )
    } else {
      p1_name <- input$player1
      p2_name <- input$player2
      paste(
        sprintf("%s won in %d of 1000 simulations (%.2f%%).", p1_name, result$p1_wins, (result$p1_wins / 10)),
        sprintf("%s won in %d of 1000 simulations (%.2f%%).", p2_name, result$p2_wins, (result$p2_wins / 10)),
        sep = "\n"
      )
    }
  })
  
  # Show sets table if apply
  output$match_sets <- renderTable({
    result <- simulate_match()
    
    if (is.character(result) || !input$simulate_sets) {
      return(NULL)
    }
    
    result$set_results
  })
  
  # Bar plot with results
  output$match_plot <- renderPlot({
    result <- simulate_match()
    
    if (is.character(result)) {
      return(NULL)
    }
    
    if (input$simulate_sets) {
      data_plot <- data.frame(
        Jugador = c(input$player1, input$player2),
        Sets_Ganados = c(result$p1_sets, result$p2_sets)
      )
      
      ggplot(data_plot, aes(x = Jugador, y = Sets_Ganados, fill = Jugador)) +
        geom_bar(stat = "identity") +
        labs(title = "Results by sets", y = "Sets won", x = "Player") +
        theme_minimal()
      
    } else {
      data_plot <- data.frame(
        Jugador = c(input$player1, input$player2),
        Victorias = c(result$p1_wins, result$p2_wins)
      )
      
      ggplot(data_plot, aes(x = Jugador, y = Victorias, fill = Jugador)) +
        geom_bar(stat = "identity") +
        labs(title = "Simulation results", y = "Number of victories", x = "Player") +
        theme_minimal()
    }
  })
  
  # Live Probability Chart
  output$match_evolution <- renderPlot({
    result <- simulate_match()
    
    if (is.character(result) || input$simulate_sets) {
      return(NULL)
    }
    
    ggplot(result$sim_evolution, aes(x = Simulación, y = Probabilidad, color = Jugador)) +
      geom_line() +
      labs(title = "Live probabilities", y = "Winning probability", x = "simulation") +
      theme_minimal()
  })
}

shinyApp(ui_2, server_2)

```

